### 2024-08-08 manuscripts/labour-v5/manuscript_main.tex
LABOUR Comments

The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable aspects and characteristics that are important in the labor market. Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name. Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages. Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men. Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics. These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.

@@

% Using the CPS from 1994 to 2019 and the 1960 to 2000 US censuses, I estimate the effect of having a Hispanic last name on labor market outcomes. In other words, I approximate the variation in labor market perception of ethnic signals, in this case having a Hispanic last name, on annual log earnings. I also study the effect of identifying as Hispanic on labor market outcomes.

% I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences. I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name. The gap becomes statistically insignificant after controlling for education. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.

\section*{Data}\label{sec:data}

@@

\section*{Empirical Strategy}\label{sec:emp_model}

@@

Using the ``synthetic'' parents sample that I constructed using the Census data, I examine the family background of the different type of children. WW couples have higher education, 12.58 years for husbands and 12.36 for wives. As a household, WW couples have 24.95 years of schooling. Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling. As a household, HH couples have 17.13 years of education. As predicted, the inter-ethnic couples exhibit matching in which partners marry people similar to them. WH husbands have 11.82 years of education, while wives have an average of 10.71 years. WH household attained a total of 22.68 years of schooling in total. HW husbands have 10.33 years of education, while wives have an average of 11.01 years. HW household attained a total of 21.50 years of schooling in total. Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple. More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers. Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education.

@@

\section*{Results}\label{sec:results}

@@

\section*{Conclusion}\label{sec:con1}

@@

While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination. Education itself can be influenced by bias, potentially resulting in divergent outcomes. This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more. Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.

>>>>

The main identifying assumption of my empirical strategy depends on the assumption that people born to HW parents are similar to their WH peer in all observable---and unobservable---aspects and characteristics that are important in the labor market. Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name. Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages. Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men. Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics. These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.

@@

Moreover, a large literature of audit studies in education economics investigated the effect of racial and ethnic signal on access to education. \textcite{bergman2018education} found that students with Hispanic sounding names received 2 percentage points fewer responses from schools than students with White sounding names. \textcite{janssen2022guidance} found that guidance counselors restricted Asian students from advanced opportunities. \textcite{gaddis2024racial} found significant discrimination in interactions with principals against Hispanic and Chinese American families. Finally, \textcite{bourabain2023school} found more evidence of discrimination in access to education against students with underprivileged backgrounds in the Flemish education system. This paper contributes to this literature by providing evidence that discrimination in access to education could lead to lower earnings for Hispanics.

The rest of the paper is organized as follows. In Section \ref{sec:data}, I describe the data used in this paper. In Section \ref{sec:emp_model}, I present the empirical strategy. In Section \ref{sec:results}, I present the results from the estimation of the two specifications. Finally, in Section \ref{sec:con1}, I conclude.

\section{Data}\label{sec:data}

@@

\section{Empirical Strategy}\label{sec:emp_model}

@@

Using the ``synthetic'' parents sample that I constructed using the Census data, I examine the family background of the different type of children. WW couples have higher education, 12.58 years for husbands and 12.36 for wives. As a household, WW couples have 24.95 years of schooling. Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling. As a household, HH couples have 17.13 years of education. As predicted, the inter-ethnic couples exhibit matching in which partners marry people similar to them. WH husbands have 11.82 years of education, while wives have an average of 10.71 years. WH household attained a total of 22.68 years of schooling in total. HW husbands have 10.33 years of education, while wives have an average of 11.01 years. HW household attained a total of 21.50 years of schooling in total. Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple. More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers. Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education \autocite{gould2020does}.

@@

\section{Results}\label{sec:results}

@@

There is a large literature in education economics that shows disparities in access to education. In an audit study of charter and traditional public schools, \textcite{bergman2018education,gaddis2024racial} find that students with Hispanic names---compared to students that are presumed White---are less likely to get a response to inquiries from schools. This could be an indication that people with Hispanic sounding names could end up lower access to education, including high value-added schools. Therefore, the earnings gap between people with Hispanic sounding last names and those with native sounding last names could be due to differences in access to education that are in turn due to discrimination.

\section{Conclusion}\label{sec:con1}

@@

While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination. Education itself can be influenced by bias, potentially resulting in divergent outcomes \autocite{bergman2018education,gaddis2024racial}. This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more. Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.

### 2024-08-08 manuscripts/labour-v5/manuscript_title_etc.tex
LABOUR Comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 5. Define Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\PAPERABSTRACT}{
Do individuals with ethnically Hispanic names face labor market discrimination? In this study, I analyze the impact of Hispanic-sounding surnames on wages, considering inter-ethnic children and the distinct traits of native and Hispanic surnames. I examine the earnings of individuals with one White and one Hispanic parent, finding that those with Hispanic surnames often experience wage disparities. My findings reveal a notable wage gap favoring individuals with White surnames. Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this difference can be attributed to educational variances. However, this shouldn't imply an absence of discrimination; it may reflect discrimination in human capital accumulation. Furthermore, I investigate the impact of self-identifying as Hispanic on earnings. Men with a Spanish-sounding last name who identify as Hispanic earn notably less than those who do not, largely due to educational differences.
\PAPERJEL}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% 5. Define Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\PAPERABSTRACT}{
% Do individuals with ethnically Hispanic names face labor market discrimination? In this study, I analyze the impact of Hispanic-sounding surnames on wages, considering inter-ethnic children and the distinct traits of native and Hispanic surnames. I examine the earnings of individuals with one White and one Hispanic parent, finding that those with Hispanic surnames often experience wage disparities. My findings reveal a notable wage gap favoring individuals with White surnames. Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this difference can be attributed to educational variances. However, this shouldn't imply an absence of discrimination; it may reflect discrimination in human capital accumulation. Furthermore, I investigate the impact of self-identifying as Hispanic on earnings. Men with a Spanish-sounding last name who identify as Hispanic earn notably less than those who do not, largely due to educational differences.

Do individuals with Hispanic names face labor market discrimination? This study analyzes the impact of Hispanic-sounding surnames on wages among inter-ethnic children with one White and one Hispanic parent. I find that individuals with Hispanic surnames often earn less, with a notable wage gap favoring those with White surnames. Males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, largely due to educational differences. Additionally, men with Spanish-sounding last names who identify as Hispanic earn significantly less, again due to educational disparities. \PAPERJEL}
